Article Title

### 2013-10-29 sections/Figure5.tex
added Figure 6.
refined Alex's revision of the results sections -- to be continued.

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure5.pdf}
\caption{
Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
{\bf A:} The correlation matrix from Fig.~\ref{fig:01}D. 
{\bf B:} Optimal cross-validated sparse+low-rank estimate of the corrleation matrix from the same dataset. 
{\bf C:} Scaled negative inverse of the correlation matrix in {\bf A} (partial pairwise correlations). 
{\bf D:} Negative inverse of the sparse+low-rank estimate in {\bf B}. 
{\bf E:} The sparse component of {\bf D} with 96.5\% off-diagonal coefficients set to zero.
{\bf F:} The low-rank component of {\bf D}, rank=7.
{\bf G:} The spatial arrangement of cells used in the analysis, connected by edges that correspond to the non-zero elements of the sparse component in {\bf E}. Green edges denote positive weights, red edges denote negative weights.  Color-filled cells are significantly tuned to grating orientation as indicated by the color code in the legend.
}
\label{fig:05}
\end{figure}

>>>>

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure5.pdf}
\caption{
Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
{\bf A:} The correlation matrix from Fig.~\ref{fig:01}D. 
{\bf B:} Partial pairwise correlations (scaled negative inverse of the correlation matrix in {\bf A}) 
{\bf C:} Optimal cross-validated sparse+latent estimate of the corrleation matrix from the same dataset. 
{\bf D:} Negative inverse of the sparse+latent estimate in {\bf C}. 
{\bf E:} The sparse component of {\bf D} with 96.5\% off-diagonal coefficients set to zero with the average of 5.1 edges per cell.
{\bf F:} The low-rank component of {\bf D}, rank=7.
{\bf G:} The spatial arrangement of cells used in the analysis, connected by edges that correspond to the non-zero elements of the sparse component in {\bf E}. Green edges denote positive weights, red edges denote negative weights.  Color-filled cells are significantly tuned to grating orientation as indicated by the color code in the legend.
}
\label{fig:05}
\end{figure}

### 2013-10-29 sections/Results.tex
refined Alex's revision of the results sections -- to be continued.

\subsection*{Covariance estimation}

Our goal is to estimate the covariance matrix, $\Sigma$, of a population of $n$ neurons. For small populations or in the limit of infinite data this can be done in a straightforward manner by using the sample covariance estimator
\begin{equation}
\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
\end{equation}
where $x(t)$ are sequential observations of population activity and $\nu$ is the number of degrees of freedom ($\nu=n-1$ if observations are independent).

However, for large population sizes and limited data this approach becomes ineffective at capturing the true structure of the covariance matrix (REFS). The number of parameters to be estimated grows quadratically with the population size. Thus, to obtain a "good" estimate of the true covariance, we need to rely on regularization -- the deliberate biasing of the estimate toward a low-dimensional, less variable target estimate \cite{Bickel:2006,Ledoit:2004}. Various such targets exist, each expressing a different idea about what a parsimonious model would look like.

We considered four different estimators, based on a few simple concepts for regularizing the covariance: independence, low-rank structure and sparse interactions. We optimized all four estimators with respect to Gaussian log-likelihood (see Discussion for a more in-depth justification of this loss function) and used cross-validation to determine hyper-parameters.

The first and probably simplest regularized estimator is shrinkage towards a diagonal matrix
\begin{equation}
\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda D,
\end{equation}
where $D$ is a diagonal matrix and the mixing proportion, $\lambda$, is determined by cross-validation. This estimator expresses the idea that in the absence of evidence for strong correlations between cells we assume they are independent (Fig.~1A). This approach is sometimes also referred to as ridge regression.

The second estimator we consider is shrinkage towards a low-rank matrix
\begin{equation}
\hat \Sigma_{\rm low-rank} = (1-\lambda) \hat \Sigma_0 + \lambda T,
\end{equation}
where $T = LL^\T + \Psi$ is a factor analysis model (REFS) consisting of a low-rank component $L$ with rank $d < n$ and a diagonal matrix $\Psi$. This estimator expresses the idea that the fluctuations in population activity are driven by a small number of unobserved sources that affect many cells (Fig.~1B).

>>>>

\subsection*{Covariance estimation}
We aim to estimate the true covariance matrix 
\begin{equation}
\Sigma = \E{(x-\mu)(x-\mu)^\T}
\end{equation}
where $\E{\cdot}$ denotes expectation  and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.

The usual estimator of the covariance matrix is the \emph{sample covariance matrix} computed from the empirical sample  $x(1),\ldots,x(n)$:
\begin{equation}
\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
\end{equation}
where $x(t)$ are sequential observations of population activity and $\nu$ is the number of degrees of freedom ($\nu=n-1$ if observations are independent).

The sample covariance matrix is unbiased i.e.\;$\E{\hat\Sigma_0}-\Sigma=0$.  For finite sample sizes, however, $\hat\Sigma_0$ is not as close to $\Sigma$ as a number of biased estimators that rely on \emph{regularization}. Regularization is the deliberate biasing of the estimate toward a low-dimensional, less variable \emph{target estimate} to strike a favorable balance between bias and variability of the estimate \cite{Bickel:2006,Ledoit:2004}.  The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators.

We considered four regularized estimators based on distinct families of low-dimensional target estimates: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors. These target estimates are depicted graphically in Fig.~\ref{fig:02}.

\input{sections/Figure2.tex}

The relative performance of the four estimators on synthetic and empirical data was evaluated by cross-validated multivariate normal log-likelihood (see Methods for detailed justification).  
Each estimator has one or two \emph{hyperparameters}, which are estimated within training datasets by nested cross-validation (See Methods).

In the first regularized estimator, the target estimate is a diagonal matrix $\hat D$.  The optimized estimate $\hat\Sigma_{\rm diag}$ is obtained by the linear \emph{shrinkage} of the unbiased estimate $\hat\Sigma_0$ toward $\hat D$ as regulated by scalar \emph{shrinkage intensity} $\lambda \in [0, 1]$:
\begin{equation}
\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda \hat D
\end{equation}
The diagonal target estimate expresses the idea of lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}A).  If this assumption aptly describes recorded data, then strong shrinkage toward $\hat D$ will add little bias while strongly reducing the variability of the estimate.

In the second regularized estimator, the target estimate is the factor model $\hat F = LL^\T + \Psi$ with $d$ factors so that $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the indvidual variances of each neuron.
Then the estimate is 
\begin{equation}
\hat \Sigma_{\rm factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
\end{equation}
This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $\hat F$ expresses the idea that the fluctuations in population activity must be driven by relatively small number of unobserved sources that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}B).

### 2013-10-29 sections/Theory.tex
refined Alex's revision of the results sections -- to be continued.

\subsection*{The sample covariance matrix}
We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons. Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.

For detailed explanation of the notation, definitions, and derivations, see Appendix.

The usual estimator of $\Sigma$ is the sample covariance matrix
\begin{equation}
\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
\end{equation}
where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom. For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom. When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.

The sample covariance matrix is constructed to be unbiased, so that $\E{\hat\Sigma_0} = \Sigma$.

\subsection*{Evaluation of covariance matrix estimators}
The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.  \Kcomment{Could we say that it quantifies the error in the estimate?}

@@

The aim of our project is to produce covariance matrix estimates that minimize the expected loss 
\begin{equation}
r = \E{\loss{\hat\Sigma, \Sigma}}
\end{equation}
which is known as the \emph{risk} of $\hat\Sigma$.

>>>>

\subsection*{Evaluation of covariance matrix estimators}
The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  The loss function quantifies the error of the estimate, \emph{i.e}.~the deviation of $\hat\Sigma$ from $\Sigma$. The loss function must attain its minimum  when $\hat\Sigma = \Sigma$.

@@

The aim of our project is to identify which covariance matrix estimator minimizes the expected loss 
\begin{equation}
r = \E{\loss{\hat\Sigma, \Sigma}}
\end{equation}
which is known as the \emph{risk} of $\hat\Sigma$.
